---
title: "HOUSEHOLD EXPENDITURE"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---
```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
library(gt)
library(ggplot2)
library(plotly)
library(tidyverse)
```

```{r}
# Importing Dataset
data1 <- read_csv("C:\\Users\\LENOVO\\OneDrive\\FILES\\PROGRAMS\\_5R\\DashBoard\\Project\\Maindataset.csv",show_col_types = FALSE)
```


1 DATASET
=======
```{r}
datatable(data1, options = list(lengthMenu = c(25, 50, 100, 500)))

```


2 CALCULATED DATA
===============
```{r}
# Calculate additional values
data1$Average_Income <- rowMeans(data1[, c('Jan Income', 'Feb Income', 'Mar Income', 'Apr Income', 'May Income', 'Jun Income')])
data1$Total_Income <- rowSums(data1[, c('Jan Income', 'Feb Income', 'Mar Income', 'Apr Income', 'May Income', 'Jun Income')])
data1$Total_Expenditure <- rowSums(data1[, c('Jan Expenditure', 'Feb Expenditure', 'Mar Expenditure', 'Apr Expenditure', 'May Expenditure', 'Jun Expenditure')])
data1$Saving <- data1$Total_Income - data1$Total_Expenditure


# Display calculated data
datatable(data1[c('Case ID', 'Economic Position', 'Occupation Type', 'Home', 'Gender', 'Region', 'Jan Income', 'Feb Income', 'Mar Income', 'Apr Income', 'May Income', 'Jun Income', 'Jan Expenditure', 'Feb Expenditure', 'Mar Expenditure', 'Apr Expenditure', 'May Expenditure', 'Jun Expenditure', 'Average_Income', 'Total_Income', 'Total_Expenditure', 'Saving')],
  options = list(lengthMenu = c(25, 50, 100, 500), pageLength = 25))
```


3 ECONOMIC POSITION
=================
```{r}
# Count occurrences of each unique value in the "Economic Position" column
economic_position_count <- table(data1$`Economic Position`)

# Create a data frame with Economic Position and ID columns
economic_position_data <- data.frame(
  Economic_Position = names(economic_position_count),
  Count_of_Persons = as.vector(economic_position_count)
)

# Count occurrences of each unique value in the "Economic Position" column
economic_position_count <- table(data1$`Economic Position`)

# Create a data frame with Economic Position and ID columns
economic_position_data <- data.frame(
  Economic_Position = names(economic_position_count),
  Count_of_Persons = as.vector(economic_position_count)
)

# Calculate grand total
grand_total <- sum(economic_position_data$Count_of_Persons)

# Add conclusions
conclusions <- data.frame(
  Conclusion = c(
    "Total count of Economic Position is :",
    "Highest position is Full-Time Work:",
    "Lowest position is Training Program:"
  ),
  Value = c(
    grand_total,
    max(economic_position_data$Count_of_Persons),
    min(economic_position_data$Count_of_Persons)
  )
)

# Create the Plotly chart for economic position distribution with data labels
plotly_chart <- plot_ly(economic_position_data, x = ~Economic_Position, y = ~Count_of_Persons, type = "bar", marker = list(color = "springgreen", width = 0.5), text = ~Count_of_Persons, hoverinfo = "text") %>%
  layout(title = "Economic Position Distribution",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Count of Persons"))

```

Column {data-width=5000}
----------------------------------------
### Chart Showing economic position distribution with data labels
```{r}
# Show the Plotly chart
plotly_chart
```

Column {data-width=5000}
----------------------------------------
### Calculated Table
```{r}
# Display Economic Position data
gt(economic_position_data)
```

### Conclusion
```{r}
# Display conclusions
gt(conclusions)
```




4_Average Income _ Toatl Income
================================
Column {data-width=5000}
-------------------------
### Average Income & Total Income
```{r}
# Calculate average income and total income by economic position
average_total_income <- data1 %>%
  group_by(`Economic Position`) %>%
  summarise(
    Average_Income = mean(`Total_Income`, na.rm = TRUE),
    Total_Income = sum(`Total_Income`, na.rm = TRUE)
  ) %>%
  bind_rows(summarise(., `Economic Position` = "Grand Total", Average_Income = mean(Average_Income), Total_Income = sum(Total_Income)))

# Display the table for average income and total income
gt(average_total_income)
```
### Conclusion
```{r}
# Create the conclusion table
conclusion <- data.frame(
  Category = c("Grand Total of Average Income", "Grand Total of Total Income",
               "Highest Average income is", "Highest Total income is",
               "Lowest Average income is", "Lowest Total income is"),
  Value = c(mean(average_total_income$Average_Income), sum(average_total_income$Total_Income),
            max(data1$Average_Income), max(data1$Total_Income),
            min(data1$Average_Income), min(data1$Total_Income))
)

# Add row numbers
conclusion <- cbind(Row = 1:nrow(conclusion), conclusion)

# Insert empty rows after the first two rows
conclusion <- rbind(conclusion[1:2, ],
                    data.frame(Row = " ", Category = " ", Value = " "),
                    conclusion[3:4, ],
                    data.frame(Row = " ", Category = " ", Value = " "),
                    conclusion[5:nrow(conclusion), ])

# Display the conclusion table
gt(conclusion)
```

Column {data-width=5000}
--------------------------
### Chart showing average income and total income by economic position
```{r}
# Filter out the "Grand Total" row
average_total_income_filtered <- average_total_income %>%
  filter(`Economic Position` != "Grand Total")

# Create the line chart
p <- average_total_income_filtered %>%
  plot_ly(x = ~`Economic Position`, 
          y = ~Average_Income,
          type = 'scatter',
          mode = 'lines',
          line = list(color = 'orange'),
          name = 'Average Income') %>%
  add_trace(y = ~Total_Income,
            type = 'scatter',
            mode = 'lines',
            line = list(color = 'skyblue'),
            name = 'Total Income') %>%
  layout(xaxis = list(title = "Economic Position"),
         yaxis = list(title = 'Income'),
         title = "Income Distribution by Economic Position")

p

```



5. Total Income _ Toatl Saving
==============================
Column {data-width=5000}
-------------------------
### Total Income & Total Saving
```{r}
# Calculate sum of total income and total saving by economic position
total_income_saving <- data1 %>%
  group_by(`Economic Position`) %>%
  summarise(
    `Total Income` = sum(Total_Income, na.rm = TRUE),
    `Total Saving` = sum(Saving, na.rm = TRUE)
  ) %>%
  bind_rows(summarise(data1, `Economic Position` = "Grand Total", `Total Income` = sum(Total_Income), `Total Saving` = sum(Saving)))

# Display the table
gt(total_income_saving)

```

### Conclusion
```{r}
# Create the conclusion table
conclusion_income_saving <- data.frame(
  Category = c("Grand Total of Total Income", "Grand Total of Total Saving",
               "Highest Income is", "Highest Saving is",
               "Lowest Income is", "Lowest Saving is"),
  Value = c(sum(total_income_saving$`Total Income`), sum(total_income_saving$`Total Saving`),
            max(total_income_saving$`Total Income`), max(total_income_saving$`Total Saving`),
            min(total_income_saving$`Total Income`), min(total_income_saving$`Total Saving`)
            )
)

# Display the conclusion table
gt(conclusion_income_saving)

```

Column {data-width=5000}
------------------------
### Chart showing total income and total saving by economic position
```{r}
# Calculate sum of total income and total saving by economic position
total_income_saving <- data1 %>%
  group_by(`Economic Position`) %>%
  summarise(
    `Total Income` = sum(Total_Income, na.rm = TRUE),
    `Total Saving` = sum(Saving, na.rm = TRUE)
  ) %>%
  bind_rows(data.frame(`Economic Position` = "Grand Total", `Total Income` = sum(data1$Total_Income, na.rm = TRUE), `Total Saving` = sum(data1$Saving, na.rm = TRUE)))

# Create the plotly chart for total income and total saving
plotly_chart_income_saving <- plot_ly(total_income_saving, x = ~`Economic Position`, y = ~`Total Income`, type = "bar", name = "Total Income", marker = list(color = "lightpink")) %>%
  add_trace(y = ~`Total Saving`, name = "Total Saving", marker = list(color = "purple")) %>%
  layout(title = "Total Income & Total Saving",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Amount"))

# Show the plotly chart
plotly_chart_income_saving

```


6.Total Income _ Toatl Expenditure
==================================
```{r}
# Calculate sum of total income and total expenditure by economic position
total_income_expenditure <- data1 %>%
  group_by(`Economic Position`) %>%
  summarise(
    `Total Income` = sum(`Total_Income`, na.rm = TRUE),
    `Total Expenditure` = sum(`Total_Expenditure`, na.rm = TRUE)
  ) %>%
  bind_rows(summarise(., `Economic Position` = "Grand Total", `Total Income` = sum(`Total Income`), `Total Expenditure` = sum(`Total Expenditure`)))


# Create the conclusion table
conclusion_income_expenditure <- data.frame(
  Category = c("Grand Total of Total Income", "Grand Total of Total Expenditure",
               "Highest Total Income is", "Highest Total Expenditure is",
               "Lowest Total Income is", "Lowest Total Expenditure is"),
  Value = c(sum(data1$`Total_Income`, na.rm = TRUE), sum(data1$`Total_Expenditure`, na.rm = TRUE),
            max(data1$`Total_Income`, na.rm = TRUE), max(data1$`Total_Expenditure`, na.rm = TRUE),
            min(data1$`Total_Income`, na.rm = TRUE), min(data1$`Total_Expenditure`, na.rm = TRUE))
)


# Create a Plotly grouped bar chart with data labels inside the bars
plotly_chart_income_expenditure <- plot_ly(total_income_expenditure, x = ~`Economic Position`, type = "bar", text = ~paste("Total Income:", round(`Total Income`, 2)), hoverinfo = "text") %>%
  add_trace(y = ~`Total Income`, name = "Total Income", marker = list(color = "red"), textposition = "inside", textfont = list(color = "white", size = 12)) %>%
  add_trace(y = ~`Total Expenditure`, name = "Total Expenditure", marker = list(color = "yellow"), text = ~paste("Total Expenditure:", round(`Total Expenditure`, 2)), textposition = "inside", textfont = list(color = "white", size = 12)) %>%
  layout(title = "Total Income & Total Expenditure",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Amount"))
```
Column {data-width=5000}
--------------------------
### Chart showing total income and total expenditure by economic position
```{r}
# Show the Plotly grouped bar chart
plotly_chart_income_expenditure
```

Column {data-width=5000}
--------------------------
### Total Income & Total Expenditure
```{r}
# Display the table
gt(total_income_expenditure)
```
### Conclusion
```{r}
# Display the conclusion table
gt(conclusion_income_expenditure)
```



DASHBOARD
==========

Row
-----

### **Economic Position**

```{r}
# Convert the calculated table to a Plotly bar chart with adjusted bar thickness
plotly_chart <- plot_ly(economic_position_data, x = ~Economic_Position, y = ~Count_of_Persons, type = "bar", marker = list(color = "blue", width = 0.5)) %>%
  layout(title = "Economic Position Distribution",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Count of Persons"))

# Show the Plotly chart
plotly_chart

```


### **Average Income and Total Income Per Economic Position**

```{r}
# Filter out the "Grand Total" row
average_total_income_filtered <- average_total_income %>%
  filter(`Economic Position` != "Grand Total")

# Create a Plotly area chart with data labels
p <- average_total_income_filtered %>%
  plot_ly(x = ~`Economic Position`, 
          y = ~Average_Income,
          type = 'scatter',
          mode = 'lines+markers',
          fill = 'tozeroy',
          line = list(color = 'yellow'),
          name = 'Average Income',
          text = ~paste("Average Income:", round(Average_Income, 2)),
          hoverinfo = "text") %>%
  add_trace(y = ~Total_Income,
            type = 'scatter',
            mode = 'lines+markers',
            fill = 'tonexty',
            line = list(color = 'skyblue'),
            name = 'Total Income',
            text = ~paste("Total Income:", round(Total_Income, 2)),
            hoverinfo = "text") %>%
  layout(xaxis = list(title = "Economic Position"),
         yaxis = list(title = 'Income'),
         title = "Income Distribution by Economic Position")

p
```

Row
----

###

```{r}

valueBox(paste("     Household Expenditure"),color = "springgreen")

```

### **Total Income and Total Saving per Economic Position**

```{r}
plotly_chart_income_saving <- plot_ly(total_income_saving, x = ~`Economic Position`, y = ~`Total Income`, type = "scatter", mode = "lines+markers", marker = list(size = 10, color = "blue"), name = "Total Income", text = ~paste("Total Income:", round(`Total Income`, 2)), hoverinfo = "text") %>%
  add_trace(y = ~`Total Saving`, marker = list(size = 10, color = "red"), name = "Total Saving", text = ~paste("Total Saving:", round(`Total Saving`, 2)), hoverinfo = "text") %>%
  layout(title = "Total Income & Total Saving",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Amount"))

# Show the Plotly line chart
plotly_chart_income_saving
```


Row
----

### **Total Income and Total Expenditure**

```{r}
# Create a Plotly grouped bar chart with data labels
plotly_chart_income_expenditure <- plot_ly(total_income_expenditure, x = ~`Economic Position`, y = ~`Total Income`, type = "bar", name = "Total Income", marker = list(color = "orange"), text = ~paste("Total Income:", round(`Total Income`, 2)), hoverinfo = "text") %>%
  add_trace(y = ~`Total Expenditure`, name = "Total Expenditure", marker = list(color = "green"), text = ~paste("Total Expenditure:", round(`Total Expenditure`, 2)), hoverinfo = "text") %>%
  layout(title = "Total Income & Total Expenditure",
         xaxis = list(title = "Economic Position"),
         yaxis = list(title = "Amount"))

# Show the Plotly grouped bar chart
plotly_chart_income_expenditure
```





